The quench action approach in finite integrable spin chains

### 2016-01-25 QAMC_draft/referee_response/response.txt

Dear editor,

We are submitting a revised version of our manuscript "The quench action approach 
in finite integrable spin chains" in which we address all the points raised 
by the referee.

Sincerely,

Vincenzo Alba
Pasquale Calabrese

---------------------------------------------------------------------
RESPONSE TO THE REFEREE
---------------------------------------------------------------------

REFEREE:

This paper represents a numerical implementation of the quench action approach,
introduced in [Phys. Rev. Lett. 110, 257203 (2013)], and in particular
addresses some questions raised in [J. Stat. Mech. (2014) P05004, Phys.
Rev. Lett. 113, 117202 (2014), J. Stat. Mech. (2014) P12009], concerning
the applicability of the method to interacting integrable models (where the
scattering matrix allows for the creation of bound states). These questions
are answered in full details and the developed numerical algorithm can be
generalized to other studies of the XXZ model in- and out-of-equilibrium.

RESPONSE:

	We thank the referee for the appreciating our work and for the positive 
	evaluation.

REFEREE:

	Abstract: There is probably a typo: ”charged” should indeed be ”charges”.

RESPONSE:

	We corrected the typo in the new version of the manuscript.

REFEREE:

	After formula 2: mutually conserved charges may be replaced with
	”mutually conserved local (and quasi-local) charges” which are the only
	one relevant for the application of the GGE logic.

RESPONSE:

	We modified the sentence according to the referee suggestion.

REFEREE:

	Page 4 : ” Physically, zero-momentum strings correspond to eigenstates
	amplitudes containing multi-particle bound states which are fully delocalized
	in the chain”. All the strings in each eigenstate are fully delocalized
	being eigenstate of the momentum operator. One may say that
	the zero momentum ones correspond to bound states of spins with zero
	propagation velocity (the post-quench light-cone velocity associated to
	them is zero).

RESPONSE:

	We thank the referee for the useful suggestion. We rephrased the sentence 
	in the new version of the manuscript to meet the referee suggestion.

REFEREE:

	Page 6 : ”the string deviations, which typically, i.e., for most of the
	chain eigenstates, vanish exponentially with L in the thermodynamic
	limit”. One may add that the complications arise when this condition
	is not valid, namely for zero momentum strings, when the deviation
	decay as power laws.

RESPONSE:

	We would like to thank the referee for the useful suggestion. We now 
	mention in the manuscript that zero-momentum strings exhibit deviations 
	which decay with a power law.

REFEREE:

	Page 8 : the observable ${\mathcal O}$ in formula 20 must be local for its matrix
	elements to be described by only a set of macroscopic distributions.

RESPONSE:

	We now rephrased the sentence as

		Moreover, for all the local observables ${\mathcal O}$ (the ones considered here)
		the contribution of the ...

	to stress the importance of the locality or quasilocality of ${\mathcal O}$

REFEREE:

	Formula 21: please define B (it is indeed defined later but here
	appears for the first time).

RESPONSE:

	We now define the Newton binomial after formula 21.

REFEREE:

	Formula 57: better use the same notation as in the determinant for the 
	string centers ($\lambda_{n;j}) since j is already used for the complex 
	rapidities)

RESPONSE:

	We modified the notation as suggested by the referee in formula 57 and 
	formula 59 as well.

REFEREE:

	After formula 71: One may comment that the vanishing of the overlaps
	with zero momentum even strings is related to the conservation laws
	of quasi-local charges [Phys. Rev. Lett. 115, 157201 (2015)]

RESPONSE:

	We now added the sentence after formula 71 as

		"This is because parity-invariant eigenstates with a single zero-momentum even-length ^M
		string, which are included in~\eref{zNeel1}, have identically zero N\'eel ^M
		overlap~\cite{brockmann-2014}. This is not related to the symmetries of the N\'eel ^M
		state, but to an ``accidental'' vanishing of the prefactor in the overlap ^M
		formula~\eref{Neel-ov}. In particular, this is related to the conservation of quasi-local
		charges~\cite{ilievski-2015a}."

	to mention the relationship between vanishing of the overlaps with zero-momentum strings 
	and conservation of quasi-local charges.

REFEREE:

	section 5.2 contains repetitions and numerous overlaps with section 2.4,
	maybe a rearrangement of the content of the two sections is necessary?

RESPONSE:

	We understand the point of the referee. However, we believe that having 
	section 5.2 somehow self-contained could be useful to the readers who are 
	numerically oriented and not interested in the Bethe ansatz details 
	of the result. Therefore, we would prefer, if the referee does not have 
	further objections, to keep section 5.2 as it is.

REFEREE:

	Monte Carlo algorithm: The authors may explain why one cannot simply
	restrict to the sector with N1 = 0 and then re-weight the whole
	sum. Indeed in reference [Phys. Rev. Lett. 110, 257203 (2013)] the
	representative state was chosen to have N1 = 0, as it was for the
	amount of zero momentum strings, therefore the logic should work in
	the same way.

RESPONSE:

	We thank the referee for this useful comment. We agree with the referee 
	that it should be possible to obtain the same results restricting to 
	the sector with N1=0. However, in our work we tried to present a numerical 
	procedure that was as general as possible. In doing so we decided to 
	include in our treatment all the spin sectors, including N1\ne0. The fact 
	that only N1=0 is the only relevant one in the thermodynamic limit 
	is not a necessary ingredient of the algorithm. Moreover, restricting to 
	N1=0 should not provide a significant improvement of the method as 
	the majority of the eigenstates of the XXX chain are in the sector with 
	N1=0;

REFEREE:

	Also would be very nice to have a plot of convergence of the Monte
	Carlo as a function of the Nsteps. One would expect that for larger
	system sizes the number of steps required is smaller if the initial state
	is chosen to be very close to the representative state. How the initial
	state of the algorithm is chosen? Did the authors try to select a proper
	discretization of the thermodynamic saddle point state ? (as done for
	example in [Phys. Rev. A 89, 033605 (2014), J. Phys. A: Math. Theor.
	48, 43FT01 (2015)])

RESPONSE:

	We thank the referee for raising these very important points. First, 
	the chosen initial state is not, in principle, important for the method. 
	The eigenstates sampled at large Monte Carlo times should not be affected 
	by the choice of the initial state. Of course a careful choice of the  
	initial state could could speed up the equilibration of the Monte Carlo. 
	However, we did not perform a sistematic study of that, and for simplicity, 
	we started the Monte Carlo simulations from the ground state of the XXX 
	chain.

	Regarding the possibility of using as initial state the discretization of 
	the saddle point state, this could certainly provide a way to improve 
	the algorithm. However, in our study we preferred not to exploit this 
	possibility in order to provide an unbiased benchmark of our numerical 
	method.

	Finally, regarding the convergence of the Monte Carlo procedure as a 
	function of the number of Monte Carlo steps, the behavior is more 
	complicated than the one described by the referee. First, in our Metropolis 
	step (formula (82) in the manuscript) we chose the proposed new eigenstate 
	|\lambda_j'\rangle completely random. This implies that in general 
	one has

		Re(\epilon'-\epsilon) \sim O(L)

	In turn this implies that if an eigenstate |\lambda_j\rangle corresponds to a 
	local minimum (not the global  one) it can be ''difficult'' for 
	the algorithm to get to a different eigenstate, because the Metropolis 
	transition probability decays exponentially as the system size increases.

	The main consequence is that at fixed number of Monte Carlo steps 
	the error bar increases dramatically upon increasing chain size. This is the 
	reason why we provide only results for L\le 60.

	This problem could be solved by changing the update scheme. The key idea would 
	be to provide in the Metropolis step a proposed new eigenstate |\lambda_j'
	\rangle such that

		\epsilon'-\epsilon \sim O(1)

	This can be clearly achieved if the Bethe number configuration corresponding to 
	|\lambda'_j\rangle are obtained from that of |\lambda_j\rangle by change only 
	``few'' Bethe numbers.

	However, the non triviality of the string content of the generic eigenstates 
	of the XXX chain, makes the implementation of this tricky, and we decided no to 
	do that in the present version of the code.

REFEREE:

	Formula 84 is generically wrong since a generic observable is a more
	involved function of the different string contents of the state (i.e. a
	determinant) and not simply a sum (which is true for the local and
	quasi-local conserved quantities). The authors should then rephrase
	then this point.

RESPONSE:

	We agree with the referee about this point. Below formula 84 we now added 
	the sentence

		"We should stress again that~\eref{ref-form} is true only
		for local or quasi-local observables, while generic observables are more complicated
		functions of the rapidities."

	to stress the importance of locality (and quasi-locality).

REFEREE:

	The authors should compute the expectation values of some simple
	local operators assion for the matrix elements are
	indeed well known (for example in arXiv:1108.6224).

RESPONSE:

	We thank the referee for this very useful comment. In fact, being able to calculate  
	generic correlation functions, besides being interesting on its own, would also provide 
	a more stringent check of the validity of the numerical method. However, 
	the correlation function proposed by the referee corresponds to the local energy 
	that we already provided in the manuscript.

	In order to benchmark our results for higher spins correlation functions we 
	now include in Figure 5 Monte Carlo results for the fluctuations of the local energy 
	density Q_2. Similar to Q4 this  corresponds to a four-spin opearator.

	We could also calculate higher spin correlation functions or some long range 
	correalation functions using known results (similar to 	the one mentioned by the 
	referee). However, this would imply a considerable numerical effort and we believe 
	it goes beyond the scope of the present paper.

REFEREE:

	Once implemented this last point, the ensemble of states created by the
	Monte Carlo algorithm should be enough to reproduce the whole time
	evolution after the quench with the logic of the quench action method
	(see [Phys. Rev. Lett. 110, 257203 (2013), J. Phys. A: Math. Theor.
	48, 43FT01 (2015)]). Can the authors comment on this?

RESPONSE:

	We thank the referee for raising this interesting point. We are in fact aware 
	of the possibility of simulating the full time evolution after a quantum quench 
	starting from the saddle point result. We believe that this would represent 
	a breaktrhough and could provide valuable insights to improve for instance 
	existing numerical methods such as tDMRG. We think that this would deserve a lot 
	of attention on its own and we plan to address it in future work.

### 2016-01-25 QAMC_draft/fs_qa_v9.tex
response

We consider the problem of constructing the stationary state following a quantum quench, 
using the exact overlaps for finite size integrable models. 
We focus on the isotropic Heisenberg spin chain with initial state N\'eel or Majumdar-Ghosh (dimer),
although the proposed approach is valid for an arbitrary integrable model. 
We consider only eigenstates which do not contain zero-momentum strings because the latter 
are affected by fictitious singularities that are very difficult to take into account. 
We show that the fraction of eigenstates that do not contain zero-momentum strings is vanishing 
in the thermodynamic limit. Consequently, restricting to this part of the Hilbert space leads to 
vanishing expectation values of local observables. However, it is possible to reconstruct the 
asymptotic values by properly reweighting the expectations in the considered subspace, at the 
price of introducing finite-size corrections. We also develop a Monte Carlo sampling of the 
Hilbert space which allows us to study larger systems. We accurately reconstruct the expectation 
values of the conserved charged and the root distributions in the stationary state, which turn 
out to match the exact thermodynamic results. The proposed method can be implemented even in 
cases in which an analytic thermodynamic solution is not obtainable.

@@

In integrable models, however, the presence of local or 
quasi-local integrals of motion strongly affects the dynamics, preventing the 
onset of thermal behavior. It has been suggested in Ref.~\cite{rigol-2007,
rigol-2008} that in this situation the post-quench steady-state can be described 
by the so-called Generalized Gibbs Ensemble ($GGE$) as 
%
\begin{equation}
\label{gge}
\langle{\mathcal O}\rangle_{GGE}=\frac{1}{Z}\Tr({\mathcal O}\rho^{GGE}), 
\quad\textrm{with}\quad\rho^{GGE}\equiv\frac{1}{Z}\exp\Big(-\sum_j\beta_jQ_j\Big). 
\end{equation}
%
Here $Q_j$ are mutually commuting conserved quantities, i.e., $[{\mathcal H},Q_j]=0\,
\forall j$ and $[Q_j,Q_k]=0\,\forall j,k$, whereas $\beta_j$ are Lagrange multipliers 
to be fixed by imposing that  $\langle Q_j\rangle_{GGE}=\langle\Psi_0|Q_j|\Psi_0\rangle$. 
This also provides  useful sum rules that will be of interest in this paper. 
The validity of the GGE has been confirmed in non-interacting theories~\cite{calabrese-2006,
cramer-2008,barthel-2008,rossini-2009,calabrese-2011,cazalilla-2012,se-12,mossel-2012a,
collura-2013,fagotti-2013,kcc14,kcc14a,sotiriadis-2014,bkc-14}, whereas in interacting 
ones the scenario is still not fully settled and some recent works~\cite{ilievski-2015a,
cardy-2015} suggest that the $GGE$ description of the steady state  is complete provided 
that the so-called quasi-local charges~\cite{prosen-2014,pereira-2014,ilievski-2015,essler-2015} 
are included in the GGE~\eref{gge}.

@@

We first present a detailed overview of the distribution of the overlaps between 
the initial states and the eigenstates of the model (overlap distribution function). 
Precisely, we provide numerical results for all the overlaps for finite chains up 
to $L\le 38$. Our results are obtained exploiting the analytical formulas 
presented in Ref.~\cite{brockmann-2014} and~\cite{pozsgay-2014a}. Crucially, 
we restrict ourselves to a truncated Hilbert space, considering only eigenstates 
of~\eref{xxx-ham} that do not contain zero-momentum strings. Physically, 
zero-momentum strings correspond to eigenstates amplitudes containing multi-particle  
bound states which are fully delocalized in the chain. From the Bethe ansatz perspective, 
in the thermodynamic limit the presence of zero-momentum strings leads to fictitious 
singularities in the overlap formulas. Dealing with these singularities is 
a formidable task that requires detailed knowledge of the solutions of the 
Bethe equations, and it can be done only in very simple cases, e.g., for small 
chains (as explicitly done for the attractive Lieb-Liniger gas for small number 
of particles \cite{calabrese-2014}). On the other hand, it has been argued that 
zero-momentum strings are irrelevant for the reconstruction of the representative 
state in the thermodynamic limit~\cite{wouters-2014A}.

@@

In the thermodynamic limit $L\to\infty$ the solutions of the Bethe equations~\eref{ba-eq} 
form particular ``string'' patterns in the complex plane, (string hypothesis)~\cite{
bethe-1931,taka-book}. Specifically, the rapidities forming a ``string'' of length $1
\le n\le M$ (that we defined here as $n$-string) can be parametrized as 
%
\begin{equation}
\label{str-hyp}
\lambda^{j}_{n;\gamma}=\lambda_{n;\gamma}-i(n-1-2j)+i\delta_{n;\gamma}^j,\qquad 
j=0,1,\dots, n-1, 
\end{equation}
%
with $\lambda_{n;\gamma}$ being the real part of the string (string center), 
$\gamma$ labelling strings with different centers, and $j$ labelling the different 
components of the string. In~\eref{str-hyp} $\delta_{n;\gamma}^j$ are the string 
deviations, which typically, i.e., for most of the chain eigenstates, vanish 
exponentially with $L$ in the thermodynamic limit. Note that real rapidities 
correspond to strings of unit length ($1$-strings, i.e., $n=1$ in~\eref{str-hyp}).

@@

In the thermodynamic limit $L\to\infty$ at fixed finite particle density $M/L$ 
the roots of the BGT equations~\eref{bgt-eq} become dense. One then defines 
the BGT root distributions for the $n$-strings as $\pmb{\rho}\equiv\{\rho_n(
\lambda)\}_{n=1}^\infty$, with $\rho_n(\lambda)\equiv\lim_{L\to\infty}[
\lambda_{n;\gamma+1}-\lambda_{n;\gamma}]^{-1}$. Consequently, the  BGT 
equations~\eref{bgt-eq} become an infinite set of coupled non-linear integral 
equations for the $\rho_n(\lambda)$ as 
%
\begin{equation}
\label{bgt-th}
a_n(\lambda)=\rho_n(\lambda)+\rho^h_n(\lambda)+\sum_m(T_{n,m}*\rho_m)
(\lambda),
\end{equation}
%
where $\rho_n^{h}(\lambda)$ are the so-called hole-distributions, and the functions 
$a_n(\lambda)$ are defined as 
%
\begin{equation}
a_n(x)\equiv\frac{1}{\pi}\frac{n}{x^2+n^2}. 
\end{equation}
%
In~\eref{bgt-th} $T_{n,m}*\rho_m$ denotes the convolution 
%
\begin{equation}
(T_{n,m}*\rho_m)(\lambda)\equiv\int_{-\infty}^{+\infty}T_{n,m}(\lambda-\lambda')
\rho_{m}(\lambda'),
\end{equation}
%
with the matrix $T_{n,m}(x)\equiv\Theta'(x)$ being defined as 
%
\begin{eqnarray}
\nonumber\fl T_{m,n}(x)\equiv\left\{\begin{array}{cc}
a_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2a_{|n-m|+2r}(x)
+a_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2a_{2r}(x)+
a_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
Given a generic, smooth enough, observable ${\mathcal O}$, in thermodynamic 
limit its eigenstate expectation value is replaced by a functional of the root 
densities $\pmb{\rho}$ as $\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle$. 
Moreover, for all the local observables considered here the contribution of the 
different type of strings factorize, and $\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}
\rangle$ becomes 
%
\begin{equation}
\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle=\sum_{n=1}^\infty
\int_{-\infty}^{+\infty}d\lambda \rho_n(\lambda) {\mathcal O}_n(\lambda), 
\end{equation}
% 
with ${\mathcal O}_n(\lambda)$ the $n$-string contribution to the expectation 
value of ${\mathcal O}$.

@@

We anticipate that the total number of parity-invariant eigenstates $Z_{Neel}$ 
for a chain of length $L$ is given as 
%
\begin{equation}
\label{N-count}
Z_{Neel}=2^{\frac{L}{2}-1}+\frac{1}{2}B\Big(\frac{L}{2},\frac{L}{4}\Big)+1. 
\end{equation}
%
On the other hand, after excluding the zero-momentum strings one obtains 
%
\begin{equation}
\label{N-count-nz}
\widetilde Z_{Neel}=B\Big(\frac{L}{2},\frac{L}{4}\Big). 
\end{equation}
%
Note that~\eref{N-count} is only an upper bound for the number of eigenstates 
with non-zero N\'eel overlap, while~\eref{N-count-nz} is exact. Before 
proceeding, we should stress that since the N\'eel state is not $SU(2)$ invariant, 
eigenstates with non-zero N\'eel overlap can contain infinite rapidities. Thus, 
one has to consider all the possible sectors with $L/2=\ell+N_{\infty}$, and $\ell$ 
($N_\infty$) the number of finite (infinite) rapidities. Notice that this 
is different for the Majumdar-Ghosh state, for which only the parity-invariant 
eigenstates in the sector with $\ell=L/2$ have to be considered (see below).

@@

Finally, using the string hypothesis and the parity-invariance condition, the prefactor of the 
determinant ratio in~\eref{Neel-ov} becomes  
%
\begin{equation}
\label{Neel-k}
\fl\quad\prod\limits_{j=1}^m\frac{\sqrt{\tilde\lambda_j^2+1}}{4\tilde\lambda_j}=
\frac{1}{4^m}\prod\limits_{j=1}^m\prod\limits_{\ell=1}^{\tilde s_j}\left[\frac{\sqrt{j^2+
\tilde\lambda^2_{j;\ell}}}{\tilde\lambda_{j;\ell}}
\prod\limits_{k=0}^{\lceil j/2\rceil-1}\frac{(2k)^2+\tilde\lambda^2_{j;\ell}}{(2k+1)^2+
\tilde\lambda^2_{j;\ell}}\right]^{(-1)^j}, 
\end{equation}
%
where $\tilde s_j$ is the number of $j$-string pairs in the Bethe state.

@@

The overlap between a generic eigenstate of the $XXX$ chain $|\{\pm\tilde\lambda_j\}
\rangle$ and the Majumdar-Ghosh state can be obtained from the N\'eel state 
overlap~\eref{Neel-ov} as~\cite{pozsgay-2014a} 
%
\begin{equation}
\label{mg-ov}
\langle MG|\{\pm\tilde\lambda_j\}_{j=1}^m\rangle=\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\tilde\lambda_j-i}{\tilde\lambda_j+i}\Big)
\Big(1+\frac{\tilde\lambda_j+i}{\tilde\lambda_j-i}\Big)
\langle N|\{\pm\tilde\lambda_j\}_{j=1}^m\rangle.
\end{equation}
%
Notice that the Bethe states having non-zero overlap with the Majumdar-Ghosh state 
do not contain infinite rapidities ($N_\infty=0$), in contrast with the N\'eel case 
(cf.~\eref{Neel-ov}). Using the string hypothesis, the multiplicative factor 
in~\eref{mg-ov} is rewritten as 
%
\begin{eqnarray}
\label{mg-k}
\fl\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\tilde\lambda_j-i}{\tilde\lambda_j+i}\Big)
\Big(1+\frac{\tilde\lambda_j+i}{\tilde\lambda_j-i}\Big)=\\\nonumber\qquad
2^m\prod\limits_{j=1}^m\prod\limits_{\ell=1}^{\tilde s_j}
\tilde\lambda_{j;\ell}^{1+(-1)^j}(\tilde\lambda^2_{j;\ell}+j^2)\prod\limits_{k=0}^{
\lfloor j/2\rfloor}\Big[\tilde\lambda_{j;\ell}^2+\Big(2k+\frac{1-(-1)^j}{2}\Big)^2
\Big]^{-2}.
\end{eqnarray}
%

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=.95\textwidth]{./draft_figs/full_vs_mc.pdf}
\end{center}
\caption{Energy density expectation values for a quench from N\'eel (a) and Majumdar-Ghosh state (b) obtained by means of the 
reweighting \eref{rewe}.
For the Majumdar-Ghosh state a quadratic fit (forcing the asymptotic value $Q_2^{(0)}=-5/8$) describes perfectly the reweighed 
data.
For the N\'eel state we also report the Monte Carlo data obtained in the next section, showing perfect agreement with the 
reweighed Bethe ansatz data. 
}
\label{fig:rew}
\end{figure}

>>>>

We consider the problem of constructing the stationary state following a quantum quench, 
using the exact overlaps for finite size integrable models. 
We focus on the isotropic Heisenberg spin chain with initial state N\'eel or Majumdar-Ghosh (dimer),
although the proposed approach is valid for an arbitrary integrable model. 
We consider only eigenstates which do not contain zero-momentum strings because the latter 
are affected by fictitious singularities that are very difficult to take into account. 
We show that the fraction of eigenstates that do not contain zero-momentum strings is vanishing 
in the thermodynamic limit. Consequently, restricting to this part of the Hilbert space leads to 
vanishing expectation values of local observables. However, it is possible to reconstruct the 
asymptotic values by properly reweighting the expectations in the considered subspace, at the 
price of introducing finite-size corrections. We also develop a Monte Carlo sampling of the 
Hilbert space which allows us to study larger systems. We accurately reconstruct the expectation 
values of the conserved charges and the root distributions in the stationary state, which turn 
out to match the exact thermodynamic results. The proposed method can be implemented even in 
cases in which an analytic thermodynamic solution is not obtainable.

@@

In integrable models, however, the presence of local or 
quasi-local integrals of motion strongly affects the dynamics, preventing the 
onset of thermal behavior. It has been suggested in Ref.~\cite{rigol-2007,
rigol-2008} that in this situation the post-quench steady-state can be described 
by the so-called Generalized Gibbs Ensemble ($GGE$) as 
%
\begin{equation}
\label{gge}
\langle{\mathcal O}\rangle_{GGE}=\frac{1}{Z}\Tr({\mathcal O}\rho^{GGE}), 
\quad\textrm{with}\quad\rho^{GGE}\equiv\frac{1}{Z}\exp\Big(-\sum_j\beta_jQ_j\Big). 
\end{equation}
%
Here $Q_j$ are mutually commuting conserved local (and quasi-local) charges, i.e., $[{\mathcal H},Q_j]=0\,
\forall j$ and $[Q_j,Q_k]=0\,\forall j,k$, whereas $\beta_j$ are Lagrange multipliers 
to be fixed by imposing that  $\langle Q_j\rangle_{GGE}=\langle\Psi_0|Q_j|\Psi_0\rangle$. 
This also provides  useful sum rules that will be of interest in this paper. 
The validity of the GGE has been confirmed in non-interacting theories~\cite{calabrese-2006,
cramer-2008,barthel-2008,rossini-2009,calabrese-2011,cazalilla-2012,se-12,mossel-2012a,
collura-2013,fagotti-2013,kcc14,kcc14a,sotiriadis-2014,bkc-14}, whereas in interacting 
ones the scenario is still not fully settled and some recent works~\cite{ilievski-2015a,
cardy-2015} suggest that the $GGE$ description of the steady state  is complete provided 
that the so-called quasi-local charges~\cite{prosen-2014,pereira-2014,ilievski-2015,essler-2015} 
are included in the GGE~\eref{gge}.

@@

We first present a detailed overview of the distribution of the overlaps between 
the initial states and the eigenstates of the model (overlap distribution function). 
Precisely, we provide numerical results for all the overlaps for finite chains up 
to $L\le 38$. Our results are obtained exploiting the analytical formulas 
presented in Ref.~\cite{brockmann-2014} and~\cite{pozsgay-2014a}. Crucially, 
we restrict ourselves to a truncated Hilbert space, considering only eigenstates 
of~\eref{xxx-ham} that do not contain zero-momentum strings. Physically, 
zero-momentum strings correspond to eigenstates amplitudes containing multi-particle  
bound states having zero propagation velocity. From the Bethe ansatz perspective, 
in the thermodynamic limit the presence of zero-momentum strings leads to fictitious 
singularities in the overlap formulas. Dealing with these singularities is 
a formidable task that requires detailed knowledge of the solutions of the 
Bethe equations, and it can be done only in very simple cases, e.g., for small 
chains (as explicitly done for the attractive Lieb-Liniger gas for small number 
of particles \cite{calabrese-2014}). On the other hand, it has been argued that 
zero-momentum strings are irrelevant for the reconstruction of the representative 
state in the thermodynamic limit~\cite{wouters-2014A}.

@@

In the thermodynamic limit $L\to\infty$ the solutions of the Bethe equations~\eref{ba-eq} 
form particular ``string'' patterns in the complex plane, (string hypothesis)~\cite{
bethe-1931,taka-book}. Specifically, the rapidities forming a ``string'' of length $1
\le n\le M$ (that we defined here as $n$-string) can be parametrized as 
%
\begin{equation}
\label{str-hyp}
\lambda^{j}_{n;\gamma}=\lambda_{n;\gamma}-i(n-1-2j)+i\delta_{n;\gamma}^j,\qquad 
j=0,1,\dots, n-1, 
\end{equation}
%
with $\lambda_{n;\gamma}$ being the real part of the string (string center), 
$\gamma$ labelling strings with different centers, and $j$ labelling the different 
components of the string. In~\eref{str-hyp} $\delta_{n;\gamma}^j$ are the string 
deviations, which typically, i.e., for most of the chain eigenstates, vanish 
exponentially with $L$ in the thermodynamic limit. A notable execption are the 
zero-momentum strings for which string devitions exhibit power-law decay. Note that real rapidities 
correspond to strings of unit length ($1$-strings, i.e., $n=1$ in~\eref{str-hyp}).

@@

In the thermodynamic limit $L\to\infty$ at fixed finite particle density $M/L$ 
the roots of the BGT equations~\eref{bgt-eq} become dense. One then defines 
the BGT root distributions for the $n$-strings as $\pmb{\rho}\equiv\{\rho_n(
\lambda)\}_{n=1}^\infty$, with $\rho_n(\lambda)\equiv\lim_{L\to\infty}[
\lambda_{n;\gamma+1}-\lambda_{n;\gamma}]^{-1}$. Consequently, the  BGT 
equations~\eref{bgt-eq} become an infinite set of coupled non-linear integral 
equations for the $\rho_n(\lambda)$ as 
%
\begin{equation}
\label{bgt-th}
a_n(\lambda)=\rho_n(\lambda)+\rho^h_n(\lambda)+\sum_m(T_{n,m}*\rho_m)
(\lambda),
\end{equation}
%
where $\rho_n^{h}(\lambda)$ are the so-called hole-distributions, and the functions 
$a_n(\lambda)$ are defined as 
%
\begin{equation}
a_n(x)\equiv\frac{1}{\pi}\frac{n}{x^2+n^2}. 
\end{equation}
%
In~\eref{bgt-th} $T_{n,m}*\rho_m$ denotes the convolution 
%
\begin{equation}
(T_{n,m}*\rho_m)(\lambda)\equiv\int_{-\infty}^{+\infty}T_{n,m}(\lambda-\lambda')
\rho_{m}(\lambda'),
\end{equation}
%
with the matrix $T_{n,m}(x)\equiv\Theta'(x)$ being defined as 
%
\begin{eqnarray}
\nonumber\fl T_{m,n}(x)\equiv\left\{\begin{array}{cc}
a_{|n-m|}(x)+\!\!\!\!\!\sum
\limits_{r=1}^{(n+m-|n-m|-1)/2}\!\!\!\!\!2a_{|n-m|+2r}(x)
+a_{n+m}(x) & \quad\mbox{if}
\quad n\ne m\\\fl\sum\limits_{r=1}^{n-1}2a_{2r}(x)+
a_{2n}(x) & \quad\mbox{if}\quad n=m
\end{array}\right.
\end{eqnarray}
%
Given a generic, smooth enough, observable ${\mathcal O}$, in thermodynamic 
limit its eigenstate expectation value is replaced by a functional of the root 
densities $\pmb{\rho}$ as $\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle$. 
Moreover, for all the local observables ${\mathcal O}$ (the ones considered here) 
the contribution of the 
different type of strings factorize, and $\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}
\rangle$ becomes 
%
\begin{equation}
\langle\pmb{\rho}|{\mathcal O}|\pmb{\rho}\rangle=\sum_{n=1}^\infty
\int_{-\infty}^{+\infty}d\lambda \rho_n(\lambda) {\mathcal O}_n(\lambda), 
\end{equation}
% 
with ${\mathcal O}_n(\lambda)$ the $n$-string contribution to the expectation 
value of ${\mathcal O}$.

@@

We anticipate that the total number of parity-invariant eigenstates $Z_{Neel}$ 
for a chain of length $L$ is given as 
%
\begin{equation}
\label{N-count}
Z_{Neel}=2^{\frac{L}{2}-1}+\frac{1}{2}B\Big(\frac{L}{2},\frac{L}{4}\Big)+1, 
\end{equation}
%
with $B(x,y)\equiv x!/(y!(x-y)!)$ the Newton binomial. 
On the other hand, after excluding the zero-momentum strings one obtains 
%
\begin{equation}
\label{N-count-nz}
\widetilde Z_{Neel}=B\Big(\frac{L}{2},\frac{L}{4}\Big). 
\end{equation}
%
Note that~\eref{N-count} is only an upper bound for the number of eigenstates 
with non-zero N\'eel overlap, while~\eref{N-count-nz} is exact. Before 
proceeding, we should stress that since the N\'eel state is not $SU(2)$ invariant, 
eigenstates with non-zero N\'eel overlap can contain infinite rapidities. Thus, 
one has to consider all the possible sectors with $L/2=\ell+N_{\infty}$, and $\ell$ 
($N_\infty$) the number of finite (infinite) rapidities. Notice that this 
is different for the Majumdar-Ghosh state, for which only the parity-invariant 
eigenstates in the sector with $\ell=L/2$ have to be considered (see below).

@@

Finally, using the string hypothesis and the parity-invariance condition, the prefactor of the 
determinant ratio in~\eref{Neel-ov} becomes  
%
\begin{equation}
\label{Neel-k}
\fl\quad\prod\limits_{j=1}^m\frac{\sqrt{\tilde\lambda_j^2+1}}{4\tilde\lambda_j}=
\frac{1}{4^m}\prod\limits_{n=1}^m\prod\limits_{\ell=1}^{\tilde s_n}\left[\frac{\sqrt{n^2+
\tilde\lambda^2_{n;\ell}}}{\tilde\lambda_{n;\ell}}
\prod\limits_{k=0}^{\lceil n/2\rceil-1}\frac{(2k)^2+\tilde\lambda^2_{n;\ell}}{(2k+1)^2+
\tilde\lambda^2_{n;\ell}}\right]^{(-1)^n}, 
\end{equation}
%
where $\tilde s_n$ is the number of $n$-string pairs in the Bethe state.

@@

The overlap between a generic eigenstate of the $XXX$ chain $|\{\pm\tilde\lambda_j\}
\rangle$ and the Majumdar-Ghosh state can be obtained from the N\'eel state 
overlap~\eref{Neel-ov} as~\cite{pozsgay-2014a} 
%
\begin{equation}
\label{mg-ov}
\langle MG|\{\pm\tilde\lambda_j\}_{j=1}^m\rangle=\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\tilde\lambda_j-i}{\tilde\lambda_j+i}\Big)
\Big(1+\frac{\tilde\lambda_j+i}{\tilde\lambda_j-i}\Big)
\langle N|\{\pm\tilde\lambda_j\}_{j=1}^m\rangle.
\end{equation}
%
Notice that the Bethe states having non-zero overlap with the Majumdar-Ghosh state 
do not contain infinite rapidities ($N_\infty=0$), in contrast with the N\'eel case 
(cf.~\eref{Neel-ov}). Using the string hypothesis, the multiplicative factor 
in~\eref{mg-ov} is rewritten as 
%
\begin{eqnarray}
\label{mg-k}
\fl\prod\limits_{j=1}^m\frac{1}{2}
\Big(1-\frac{\tilde\lambda_j-i}{\tilde\lambda_j+i}\Big)
\Big(1+\frac{\tilde\lambda_j+i}{\tilde\lambda_j-i}\Big)=\\\nonumber\qquad
2^m\prod\limits_{n=1}^m\prod\limits_{\ell=1}^{\tilde s_n}
\tilde\lambda_{n;\ell}^{1+(-1)^n}(\tilde\lambda^2_{n;\ell}+n^2)\prod\limits_{k=0}^{
\lfloor n/2\rfloor}\Big[\tilde\lambda_{n;\ell}^2+\Big(2k+\frac{1-(-1)^n}{2}\Big)^2
\Big]^{-2}.
\end{eqnarray}
%

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=.95\textwidth]{./draft_figs/full_vs_mc.pdf}
%\includegraphics[width=.95\textwidth]{full_vs_mc.pdf}
\end{center}
\caption{Energy density expectation values for a quench from N\'eel (a) and Majumdar-Ghosh state (b) obtained by means of the 
reweighting \eref{rewe}.
For the Majumdar-Ghosh state a quadratic fit (forcing the asymptotic value $Q_2^{(0)}=-5/8$) describes perfectly the reweighed 
data.
For the N\'eel state we also report the Monte Carlo data obtained in the next section, showing perfect agreement with the 
reweighed Bethe ansatz data. 
}
\label{fig:rew}
\end{figure}
